\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-08-28 write_up/mixfishsim_ecolmod.tex
Sharpening up text

\replaced[id = JJ]{Most fisheries}{Fishing} exploit\deleted[id=JJ]{s} a variety 
of spatially and temporally heterogeneous fish populations using
species-unselective gear that can result in unintended, unwanted catch of low
quota or protected species. Reducing these unwanted catches is crucial for
biological and economic sustainability of `mixed fisheries' and implementation
of an ecosystem approach to fishing. \\

\replaced[id = PD]{If fisheries are to avoid unwanted catch,}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions, population movement and how fishers interact with different fish
populations. This reflects the fact that data on fish location at high temporal
and spatial resolutions is expensive and difficult to collect. Proxies inferred
from either scientific surveys or commercial catch data are often used to model
distributions, usually with sparse data at limited spatial and temporal
resolution. \\

To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal simulation model
incorporating: i) delay-difference population dynamics, ii) population movement
using Gaussian Random Fields to simulate patchy, heterogeneously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based
on species targeting under an explore-exploit strategy via a mix of correlated
random walk movement (for exploration) and learned behaviour (for exploitation)
phases of the fisheries.  \\

We simulated 50 years of fishing and used the results from the fisheries catch
to draw inference on the underlying community structures. We compared this
inference to a simulated fixed-site sampling design commonly used for fisheries
monitoring purposes and the true underlying community structure. We i) used
the results to establish the potential and limitations of fishery-dependent
data in providing a robust picture of spatiotemporal distributions; and ii)
simulated an area closure based on areas defined from the different data sources
at a range of temporal and spatial resolutions to assess their effectiveness
on reducing catches of a fish population. \\

@@

\replaced[id=PD]{Adaptive spatial management strategies have}{Use of spatial
	management as a tools} been proposed as a way of reducing discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.  \deleted[id=PD]{its}
Implementation of avoidance measures is, however, restricted by lack of
knowledge of fish and fishery spatiotemporal dynamics and understanding of the
scale at which processes become important for management. Understanding the
correct scale for spatial measures is crucial for implementing solutions at a
resolution that ensures effective management \citep{Dunn2016} while minimising
economic impact.  For example, the problem can be to identify a scale that
promotes species avoidance for vulnerable or low quota species while allowing
continuance of sustainable fisheries for available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} appropriate
spatial scales for fisheries closures has been a challenge in the past but
identified as crucial to its success \citep{Costello2010, Dunn2016}, and is
hampered where spatial information is coarse due to low resolution reporting of
fisheries catch which is subject to aggregation across larger gridded areas
\citep{Branch2005}. Further, coarse spatial data may lead to poorly sited closures,
which have led to ineffectual measures with unintended consequences.  For
example, increased benthic impact on previously unexploited areas was observed
from the cod closure in the North Sea with a lack of observed intended effect
in reducing cod exploitation \citep{Rijnsdorp2001,Dinmore2003}).
\replaced[id=PD]{M}{Since then m}ore refined spatiotemporal information has
\added[id=PD]{since} become available through the combination of logbook and
Vessel Monitoring System (VMS) data \citep{Lee2010, Bastardie2010,
	Gerritsen2012, Mateo2016} and more real-time spatial management has
been possible \citep[e.g.][]{Holmes2011}.  Such information is, however,
derived from an inherently biased sampling programme, targeted fishing, where
fishers establish favoured fishing grounds through an explore-exploit strategy
\citep{Bailey2018} where they search for areas with high catches and then use
experience to return to areas where they've experienced high catch in the past.
\deleted[id=PD]{ Further, fishers generally only recorded landings (not catch)
	on a daily basis\todo{\added[id=JJ]{Its not clear what the problem is:
			landings collected on daily basis or landings recorded
			rather than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

\replaced[id=PD]{In order to understand the effect of spatiotemporal
	aggregation of data we ask two fundamental questions regarding
	inference derived from observational data:}{ and use the
	results}\deleted[id=PD]{ from the fishing model:}

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time by use of a Gaussian spatial
process to define suitable habitat\replaced[id=PD]{. Precise locations being}{
	and are} known \added[id=PD]{directly} rather than inferred from
sampling or commercial catch\added[id=PD]{, we can use the population model to
	validate how inference from fisheries-dependent and fisheries
	independent sampling relates to the real community structure in a way
	we could not with real data}. We ii) compare, at different spatial and
temporal aggregations, the 'real population' distributions to samples from
fisheries-dependent and fisheries independent catches to test if these are a
true reflection of the relative density of the populations. We then iii)
simulate a fishery closure to protect a species based on different spatial and
temporal data aggregations. We use these evaluations to draw inference on the
utility of commercial data in supporting management decisions.
\todo{\added[id=JJ]{If the paper has two goals this should be clear from the
		start, but may be better over 2 MSs}\added[id=PD]{I would like
		to keep both parts, but have made clearer in how its set out.
		The closure scenarios form validation of the data aggregation,
		rather than effectiveness of the closures themselves - so its a
		continuation of the same question in my eyes}}\deleted[id=PD]{.
	Further, we extend our analysis to a range of spatial and temporal
	scales to assess the impact of these processes on the success of the
	management measure.}\\

@@

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics (including fishing and natural
mortality which are instantaneous rates, growth of the population biomass)
operate on a daily time-step\replaced[id=PD]{, while p}{. P}opulation movement
occurs on a weekly time-step\replaced[id=PD]{. R}{, while r}ecruitment
\replaced[id=PD]{takes place}{occurs} periodically each year for a set time
\replaced[id=PD]{duration}{period (e.g. 3 weeks)} specified for each
population\deleted[id=PD]{.}\added[id=PD]{, while the fishing module operates
	on a tow-by-tow basis (i.e. multiple events a day)}.
\deleted[id=PD]{Here we describe each of the model components; 1) Population
	dynamics, 2) Recruitment dynamics, 3) Population movement dynamics, 4)
	fishery dynamics.}

>>>>

\replaced[id = JJ]{Most fisheries}{Fishing} exploit\deleted[id=JJ]{s} a variety 
of spatially and temporally heterogeneous fish populations using
species-unselective gear. This can result in unintended, unwanted catch of low
quota or protected species and reducing these catches is crucial for
biological and economic sustainability of `mixed fisheries'. \\

\replaced[id = PD]{If fisheries are to avoid unwanted catch,}{To implement
	effective spatial measures to reduce discards} a good understanding of
spatiotemporal fishery dynamics is required. However, traditional scientific
advice is limited by a lack of highly resolved knowledge of population
distributions and movement, and how fishers interact with different fish
populations. Because data on fish location at high temporal and spatial
resolutions is expensive and difficult to collect proxies are usually inferred
from scientific surveys or commercial catches with limited spatial and temporal
resolution. \\

To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal simulation model
incorporating: i) delay-difference population dynamics, ii) population movement
using Gaussian Random Fields to simulate patchy, heterogeneously distributed
populations, and iii) fishery dynamics for multiple fleet characteristics based
on species targeting under an explore-exploit strategy. This is implemented via
a mix of correlated random walk movement (for exploration) and learned
behaviour (for exploitation) phases of the fisheries.  \\

We simulated 50 years of fishing and used the results to draw inference on the
underlying community structures. We compared inference from commercial catch to
a simulated fixed-site sampling design and the true underlying community. We i)
establish the potential limitations of fishery-dependent data in providing a
robust picture of spatiotemporal distributions; and then ii) simulated an area
closure based on areas defined from the different data sources at a range of
temporal and spatial resolutions to assess their effectiveness on reducing
catches of a fish population. \\

@@

\replaced[id=PD]{Adaptive spatial management strategies have}{Use of spatial
	management as a tools} been proposed as a way of reducing overquota
discards \added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}.
\deleted[id=PD]{its} However, implementation is restricted by lack of knowledge
of fish and fishery spatiotemporal dynamics and understanding of the scale at
which processes become important for management. Understanding the correct
scale for spatial measures is crucial for implementing effective solutions
which  minimising economic impact \citep{Dunn2016}. For example, the problem
can be to identify a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} appropriate
spatial scales for fisheries closures has been a challenge in the past but is
crucial to their success \citep{Costello2010, Dunn2016}. Inference is hampered
where spatial information is coarse due to low resolution reporting of
fisheries catch which is aggregated across larger gridded areas
\citep{Branch2005}. Further, if data does not allow identification of spatial
features it may lead to poorly sited closures which are ineffectual or have
unintended consequences. For example, increased benthic impact on previously
unexploited areas from the cod closure in the North Sea were observed without
the intended effect of reducing cod exploitation
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatiotemporal information has \added[id=PD]{since} become available
through the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}. However,
fishers establish favoured fishing grounds through an explore-exploit strategy
\citep{Bailey2018} where they search for areas with high catches and then use
experience to return to areas where they've experienced high catch in the past.
This leads to an inherently biased sampling where target species are
over-represented in catch as fishers exploit areas of high abundance.
\deleted[id=PD]{ Further, fishers generally only recorded landings (not catch)
	on a daily basis\todo{\added[id=JJ]{Its not clear what the problem is:
			landings collected on daily basis or landings recorded
			rather than catch}}.  This leads to questions about the
	validity of inference that can be drawn from landings data assigned to
	VMS activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I
		thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

\replaced[id=PD]{In order to understand the effect of spatiotemporal
	aggregation of data and fishery targeting we ask two fundamental
	questions regarding inference derived from observational data:}{ and
	use the results}\deleted[id=PD]{ from the fishing model:}

@@

To answer these questions we i) develop a simulation model where population
dynamics are highly-resolved in space and time by use of a Gaussian spatial
process to define suitable habitat for different populations\replaced[id=PD]{.
	Precise locations of the fish being}{ and are} known
\added[id=PD]{directly} rather than inferred from sampling or commercial
catch\added[id=PD]{, we can use the population model to validate how inference
	from fisheries-dependent and fisheries independent sampling relates to
	the real community structure in a way we could not with real data}. We
ii) compare, at different spatial and temporal aggregations, the 'real
population' distributions to samples from fisheries-dependent and fisheries
independent catches to test if these are a true reflection of the relative
density of the populations. We then iii) simulate a fishery closure to protect
a species based on different spatial and
temporal data aggregations. \\

We use these evaluations to draw inference on the utility of commercial data in
supporting management decisions.

@@

\deleted[id=PD]{The fishing model operated on a tow-by-tow basis, while}
\replaced[id=PD]{P}{p}opulation dynamics (including fishing and natural
mortality which are instantaneous rates, and growth of the population biomass)
operate on a daily time-step\replaced[id=PD]{, while p}{. P}opulation movement
occurs on a weekly time-step\replaced[id=PD]{. R}{, while r}ecruitment
\replaced[id=PD]{takes place}{occurs} periodically each year for a set time
\replaced[id=PD]{duration}{period (e.g. 3 weeks)} specified for each
population\deleted[id=PD]{.}\added[id=PD]{, while the fishing module operates
	on a tow-by-tow basis (i.e. multiple events a day)}.
\deleted[id=PD]{Here we describe each of the model components; 1) Population
	dynamics, 2) Recruitment dynamics, 3) Population movement dynamics, 4)
	fishery dynamics.}
